### 2016-09-12 paper/conclusion.tex
added variable information on master file to each file

The growing dependence on big data makes reliably converting
between different data formats increasingly important.
\emph{Lenses} are a bi-directional technique for specifying such
conversions.  Specifically, a lens is a declarative program that
simultaneously specifies a function to convert data from type $A$ to
type $B$ and a second function to convert data from type $B$ to type
$A$.  Such programs usually provide invertibility guarantees, avoiding
pernicious bugs that can corrupt data and downstream processing.
Unfortunately, lenses can be quite difficult to write because they
require programmers to think in two directions at once.  To overcome
this difficulty, we develop an algorithm to automatically synthesize
lenses from format descriptions, specified as regular expressions, and
a small number of examples.  The rich structure of regular expressions
makes lens synthesis more complicated than previous work on other
forms of type-directed synthesis.  Specifically, finding a term to
bi-directionally map between regular expressions $A$ and $B$ requires
first finding equivalent regular expressions $A'$ and $B'$ that are
more algorithmically tractable.
We show how to manage this search by defining a new language of
lenses governed by semi-canonical regular expression types and prove it
sound with respect to a standard language of bijective lenses.  
We then show
how to search the reduced space for well-typed lenses that are consistent
with the given examples.  We demonstrate the effectiveness of our algorithm
by experimenting on a benchmark suite containing 42 examples
comprising microbenchmarks, Flash Fill
examples, and Augeas transformations of Linux configuration files.
We show we are able to infer lenses for most examples in the
benchmark suite in an average of 0.134 seconds.

%% The language of regular lenses expresses programs which can convert between
%% source and target regular languages.
%% We tackle the task of synthesizing these programs through the regular expressions
%% of their types.
%% This is difficult because we cannot use basic type directed synthesis techniques,
%% as there are many equivalences in the types of these regular expressions.
%% We do this by developing a language for partially normalized regular expressions,
%% and searching through those regular expressions for two that have a lens between
%% them which can be synthesized in a type-directed way.
%% We develop strategies for efficiently enacting this search and synthesis,
%% and evaluate our strategies against a suite of microbenchmarks and linux
%% configuration files.

>>>>

\section{Conclusion}

Conclusion

### 2016-09-12 paper/implementation.tex
added variable information on master file to each file

$\Distance_{dist}(\DNFRegex,\DNFRegexAlt)=\LOneNorm(\phi(\DNFRegex),\phi(\DNFRegexAlt))$
where \LOneNorm{} is the taxicab metric on \Module{}.
The priority of a pair of DNF regular expressions, $(\DNFRegex{},\DNFRegexAlt{})$,
is $\Distance(\DNFRegex{},\DNFRegexAlt{})+8*expansion_count$.
Keeping track of expansion count is important as it allows for lower priority
expansions to be eventually popped, instead of getting stuck doing incorrect,
increasingly deep expansions whose DNF regular expressions have a low distance.
\afm{should i do an example, or is that going into too much detail?}
Experimentally, 8 was a good value for how much weight the expansion
count should get in the priority.  It allows for not getting stuck in a local
minima in the search space, while still allowing for the distance to choose the
correct expansion the majority of the time.

@@

\begin{definition}
$\DNFOf{\Sequence_1;\ldots\Sequence_n} \leq
\DNFOf{\SequenceAlt_1;\ldots\SequenceAlt_m}$
if $n<m$ or $n=m$ and
$sort([\Sequence_1;\ldots;\Sequence_n])\leq
sort([\SequenceAlt_1;\ldots;\SequenceAlt_n])$ by the dictionary ordering
on the sequence ordering.

$\SequenceOf{\String_0;\Atom_1;\ldots;\Atom_n;\String_n} \leq
\SequenceOf{\StringAlt_0;\AtomAlt_1;\ldots;\AtomAlt_n;\String_n}$
if $n<m$ or $n=m$ and
$sort([\Atom_1\ldots\Atom_n])\leq
sort([\AtomAlt_1\ldots\AtomAlt_n])$ by the dictionary ordering on the
atom ordering.

$\RegexVariable < \IterateLens{\DNFLens}$

$\RegexVariable \leq \RegexVariableAlt$ if
$\RegexVariable \leq \RegexVariableAlt$ as strings.

$\StarOf{\DNFRegex} \leq \StarOf{\DNFRegexAlt}$ if $\DNFRegex \leq \DNFRegexAlt$.
\end{definition}

\begin{theorem}
$\DNFRegex\leq\DNFRegexAlt$ and $\DNFRegexAlt\leq\DNFRegex$ if, and only
if there exists a $\DNFLens \OfType \DNFRegex \Leftrightarrow \DNFRegexAlt$.
\end{theorem}

Beyond that, we can use these orderings to find the lenses.
If $\DNFOf{\Sequence_1;\ldots;\Sequence_n}\leq\DNFOf{\SequenceAlt_1;\ldots;\SequenceAlt_n}$,
and $\DNFOf{\SequenceAlt_1;\ldots;\SequenceAlt_n}\leq\DNFOf{\Sequence_1;\ldots;\Sequence_n}$,
then $\Sort([\Sequence_1;\ldots;\Sequence_n])\leq\Sort([\Sequence_1;\ldots;\Sequence_n])$
using the dictionary ordering, which means that
$\Sort([\Sequence_1;\ldots;\Sequence_n])[i]\leq
\Sort([\SequenceAlt_1;\ldots;\SequenceAlt_n])[i]$ and
$\Sort([\SequenceAlt_1;\ldots;\SequenceAlt_n])[i]\leq
\Sort([\Sequence_1;\ldots;\Sequence_n])[i]$, which means that there is a
sequence lens between them.
\afm{I can keep going on about this stuff, should it go here?  Should I just make a brief claim
about it and move on}

@@

>>>>

$\Distance_{dist}(\DNFRegex,\DNFRegexAlt)=\LOneNorm(\phi(\DNFRegex),\phi(\DNFRegexAlt))$
where \LOneNorm{} is the taxicab metric on \Module{}.
The priority of a pair of DNF regular expressions, $(\DNFRegex{},\DNFRegexAlt{})$,
is $\Distance(\DNFRegex{},\DNFRegexAlt{})+8*expansion_{count}$.
Keeping track of expansion count is important as it allows for lower priority
expansions to be eventually popped, instead of getting stuck doing incorrect,
increasingly deep expansions whose DNF regular expressions have a low distance.
\afm{should i do an example, or is that going into too much detail?}
Experimentally, 8 was a good value for how much weight the expansion
count should get in the priority.  It allows for not getting stuck in a local
minima in the search space, while still allowing for the distance to choose the
correct expansion the majority of the time.

@@

ATTEMPT TO WRITE IT LESS FORMALLY

Unfortunately, while synthesizing a DNF regular expression, we must synthesize
two permutations, one to deal with determining which sequence
gets mapped to which (handling the commutativity of +), and one to deal
with determining within a sequence, which atoms get mapped to which
(handling swaps).

We found an incredibly efficient strategy for doing this, by further normalizing
the DNF regular expressions.  Instead of allowing any order of sequences, we
provide a total preorder on sequences, and order order the sequences
according to that ordering.

We do the same trick for normalizing the sequences.  While reordering a sequence
changes the semantics, we can temporarily reorder the sequence, remembering the
permutation of the reordering, and recover the original sequence later.

\begin{definition}
Given an ordering $\leq : A \rightarrow A \rightarrow \SetOf{-1,0,1}$,
we denote the dictionary ordering on lists of $As$, $\ListOf{\leq}$.
\end{definition}

\begin{definition}
  Given an ordering $\leq : A \rightarrow A \rightarrow \SetOf{-1,0,1}$,
  and $\ListOf{a_1;\ldots;a_n}\in A \ListType$
  We define $\SortingOf{\leq}{\ListOf{a_1;\ldots;a_n}}$ as
  $\sigma \in \PermutationSetOf{n}$ where if $i\leq j$,
  then $a_{\sigma(i)} \leq a_{\sigma(j)}$.

  We define $\SortOf{\leq}{\ListOf{a_1;\ldots;a_n}}$ as
  $\ListOf{a_{\sigma(1)};\ldots;a_{\sigma(n)}}$ where
  $\sigma=\SortingOf{\leq}{\ListOf{a_1;\ldots;a_n}}$
\end{definition}

\begin{definition}
  \begin{enumerate}
  \item
    $\DNFOf{\Sequence_1;\ldots\Sequence_n}
    \DNFLeq
    \DNFOf{\SequenceAlt_1;\ldots\SequenceAlt_m}$
    if
    $\SortOf{\SequenceLeq}{\ListOf{\Sequence_1;\ldots;\Sequence_n}}
    \ListOf{\SequenceLeq}
    \SortOf{\SequenceLeq}{[\SequenceAlt_1;\ldots;\SequenceAlt_n]}$.

  \item
    $\SequenceOf{\String_0;\Atom_1;\ldots;\Atom_n;\String_n}
    \SequenceLeq
    \SequenceOf{\StringAlt_0;\AtomAlt_1;\ldots;\AtomAlt_n;\String_n}$
    if
    $\SortOf{\AtomLeq}{\ListOf{\Atom_1\ldots\Atom_n}}
    \ListOf{\AtomLeq}
    \SortOf{\AtomLeq}{[\AtomAlt_1\ldots\AtomAlt_n]}$.

  \item
    $\StarOf{\DNFRegex} \AtomLeq \StarOf{\DNFRegexAlt}$
    if
    $\DNFRegex \DNFLeq \DNFRegexAlt$
  \end{enumerate}
\end{definition}

This allows us to put our DNF regular expressions into a more normal form.

\begin{definition}
  \begin{enumerate}
  \item A DNF regular expression $\DNFOf{Sequence_1;\ldots;\Sequence_n}$ is
    \textit{normalized} if $i \leq j$ implies $\Sequence_1\SequenceLeq\Sequence_j$,
    and each $\Sequence_i$ is normalized.
  \item A sequence $\SequenceOf{\Atom_1;\ldots;\Atom_n}$ is \textit{normalized}
    if $i \leq j$ implies $\Atom_i\AtomLeq\Atom_j$,
    and each $\Atom_i$ is normalized.
  \item An atom $\StarOf{\DNFRegex}$ is \textit{normalized} if
    $\DNFRegex$ is normalized.
  \end{enumerate}
\end{definition}

We can also easily normalize a DNF regular expression

\begin{definition}
  \begin{enumerate}
  \item $\NormalizeOf{\DNFOf{\Sequence_1;\ldots;\Sequence_n}} =
    \DNFOf{\NormalizeOf{\Sequence_{\sigma(1)}};\ldots;\NormalizeOf{\Sequence_{\sigma(n)}}}$,
    where
    $\sigma=\SortingOf{\SequenceLeq}
    {\ListOf{\NormalizeOf{\Sequence_1};\ldots;\NormalizeOf{\Sequence_n}}}$
  \item $\NormalizeOf{\SequenceOf{\String_0;\Atom_1;\ldots;\Atom_n;\String_n}} =
    \SequenceOf{\String_0;\Atom_{\sigma(1)};\ldots;\Atom_{\sigma(n)};\String_n}$,
    where
    $\sigma=\SortingOf{\SequenceLeq}
    {\ListOf{\NormalizeOf{\Atom_1};\ldots;\NormalizeOf{\Atom_n}}}$
  \item $\NormalizeOf{\StarOf{\DNFRegex}}=\StarOf{\NormalizeOf{\DNFRegex}}$
  \end{enumerate}
\end{definition}

Critically, the normalized form of the DNF regular expression, and the ordering
on DNF regular expressions, are tied to the existance of a DNF Lens without
rewrites.

\begin{theorem}
  $\NormalizeOf{\DNFRegex}\DNFLeq\NormalizeOf{\DNFRegexAlt}$ and
  $\NormalizeOf{\DNFRegexAlt}\DNFLeq\NormalizeOf{\DNFRegex}$ if, and only
  if, there exists a $\DNFLens \OfType \DNFRegex \Leftrightarrow \DNFRegexAlt$,
  where $\DNFLens$ has no applications of a rewrite rule.
\end{theorem}

This can be proven with mutual induction, proving similar statements for
sequences and atoms.

Furthermore, two DNF regular expressions can be zipped together to form a
DNF lens.  Consider $\DNFOf{\Sequence_{\sigma_1(0)};\ldots;\Sequence_{\sigma_1(n)}}$, and
$\DNFOf{\SequenceAlt_{\sigma_2(0)};\ldots;\SequenceAlt_{\sigma_2(n)}}$, two normalized
DNF regular expressions.  $\Sequence_{\sigma(0)}$ merely needs to be mapped to
$\SequenceAlt_{\tau(0)}$.  Furthermore, the permutation mapping the two lenses
merely needs to be $\sigma_1 \Compose \sigma_2^{-1}$

We can use these orderings to find

Beyond that, we can use these orderings to find the lenses.
If $\DNFOf{\Sequence_1;\ldots;\Sequence_n}\leq\DNFOf{\SequenceAlt_1;\ldots;\SequenceAlt_n}$,
and $\DNFOf{\SequenceAlt_1;\ldots;\SequenceAlt_n}\leq\DNFOf{\Sequence_1;\ldots;\Sequence_n}$,
then $\Sorting([\Sequence_1;\ldots;\Sequence_n])\leq\Sorting([\Sequence_1;\ldots;\Sequence_n])$
using the dictionary ordering, which means that
$\Sorting([\Sequence_1;\ldots;\Sequence_n])[i]\leq
\Sorting([\SequenceAlt_1;\ldots;\SequenceAlt_n])[i]$ and
$\Sorting([\SequenceAlt_1;\ldots;\SequenceAlt_n])[i]\leq
\Sorting([\Sequence_1;\ldots;\Sequence_n])[i]$, which means that there is a
sequence lens between them.
\afm{I can keep going on about this stuff, should it go here?  Should I just make a brief claim
about it and move on}

@@

%%% Local Variables:
%%% TeX-master: "main"
%%% End:
